---
title: "Reproducible Research: Peer Assessment1"
output: html_document
---


##Loading and preprocessing the data
```{r}
data <- read.csv("./activity.csv")
```

Required libraries
```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
```

##What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
```{r}
q1Total <- data %>%
    group_by(date) %>%
    summarise(total_steps=sum(steps, na.rm=TRUE))
```

Make a histogram of the total number of steps taken each day
```{r}
hist(q1Total$total_steps, main="Total number of steps per day", xlab="Steps", ylab="Frequency")
```

Report the mean and median of the total number of steps taken per day
```{r}
summary(q1Total)
```

##What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
q2Avg <- data %>%
    group_by(interval) %>%
    summarise(avg_steps=mean(steps, na.rm=TRUE))
plot(q2Avg$interval, q2Avg$avg_steps, type="l", xlab ="interval", ylab="Average steps")
```

Calculate the maximum number of step on average across all the days in the dataset
```{r}
max_step <- max(q2Avg$avg_steps, na.rm=TRUE)
filter(q2Avg, avg_steps==max_step)
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset
```{r}
summary(data$steps)
```

Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
new_data <- data
index_of_na <- which(is.na(new_data$steps))
# replaces NA in variable 'steps' of data frame new_data with the average step value
for (i in index_of_na) {
    new_data$steps[i] <- with(q2Avg, avg_steps[interval==new_data$interval[i]])  
}
```

Make a histogram of the total number of steps taken each day 
```{r}
q3Total <- aggregate(steps~date, new_data, sum)
hist(q3Total$steps, main="Total number of steps per day", xlab="Steps", ylab="Frequency")
```

Calculate and report the mean and median total number of steps taken per day. These values differ from the estimates from the first part of the assignment
```{r}
summary(q3Total)
```

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
q4_data <- new_data
for (i in 1:length(q4_data[,2])) {
    dw <- wday(as.Date(q4_data$date[i]))
    q4_data$dayofweek[i] <- if (dw == 1 || dw == 7) {
        "weekend"
    } else {
        "weekday"
    }
}
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
q4Avg <- aggregate(q4_data$steps, list(interval=q4_data$interval, date=q4_data$dayofweek), mean)
qplot(interval, x, data=q4Avg, color=date, facets=date~., geom=c("line"),
       ylab="avgerage number of steps")
```